---
title: "01 Introduction"
output: html_notebook
---

```{r setup, include=FALSE, error=TRUE, message=FALSE, fig.asp=1}
knitr::opts_chunk$set(echo = TRUE, comment = NA, error = TRUE)
```

## Basic descriptive statistics

```{r}
library(psych)

data(sat.act)
describe(sat.act)
```

## Correlational structure
There  are  many  ways  to  display  correlations.   Tabular  displays  are  probably  the  most
common.  The output from the cor function in core R is a rectangular matrix.
lowerMat will  round  this  to  (2)  digits  and  then  display  as  a  lower  o   diagonal  matrix. lowerCor calls cor with use=`pairwise',  method=`pearson' as default values and returns (invisibly) the full correlation matrix and displays the lower o  diagonal matrix

```{r}
lowerCor(sat.act)
```

```{r fig.asp=1}
pairs.panels(sat.act, pch='.')
```
Figure 1:  Using the pairs.panels function to graphically show relationships.  The x axis
in each scatter plot represents the column variable, the y axis the row variable.  Note the
extreme outlier for the ACT. The plot character was set to a period (pch='.')  in order to
make a cleaner graph.

```{r}
female <- subset(sat.act,sat.act$gender==2)
male <- subset(sat.act,sat.act$gender==1)
lower <- lowerCor(male[-1])
```

```{r}
upper <- lowerCor(female[-1])
```

```{r}
both <- lowerUpper(lower,upper)
round(both,2)
```

It is also possible to compare two matrices by taking their differences and displaying one (below the diagonal) and the difference of the second from the first above the diagonal

```{r}
diffs <-  lowerUpper(lower,upper,diff=TRUE)
round(diffs,2)
```

## Heatmap displays of correlational structure

```{r fig.asp=1}
cor.plot(Thurstone,numbers=TRUE,main="9 cognitive variables from Thurstone")
```
Figure 2:  The structure of correlation matrix can be seen more clearly if the variables are
grouped by factor and then the correlations are shown by color.  By using the 'numbers'
option, the values are displayed as well.


```{r fig.asp=1}
circ <- sim.circ(24)
r.circ <- cor(circ)
cor.plot(r.circ,main='24 variables in a circumplex')
```
Figure 3: Using the cor.plot function to show the correlations in a circumplex.  Correlations
are highest near the diagonal, diminish to zero further from the diagonal, and the increase
again towards the corners of the matrix.  Circumplex structures are common in the study
of affect.


