---
title: "Wighted Least Square Factor Analysis"
output:
  html_document: default
  html_notebook: default
---

```{r setup, include=FALSE, error=TRUE, message=FALSE, fig.asp=1}
knitr::opts_chunk$set(echo = TRUE, comment = NA, error = TRUE)
```

## 5.1.3 Weighted Least Squares Factor Analysis
Similar to the minres approach of minimizing the squared residuals, factor method "wls"
weights the squared residuals by their uniquenesses.  This tends to produce slightly smaller 
overall residuals. In the example of weighted least squares, the output is shown by using the
print function with the cut option set to 0. That is, all loadings are shown (Table 3).

> Install psych and GPArotation packages.

```{r}
library(psych)

data(Thurstone)

f3t <- fa(Thurstone, 3, n.obs=213)
f3t
```

#### Table 2:
The 9 variable problem from Thurstone is a classic example of factoring where
there is a higher order factor, g, that accounts for the correlation between the factors.  The
extraction method was principal axis.  The transformation was a targeted transformation
to a simple cluster solution.


```{r}
f3 <- fa(Thurstone, 3, n.obs = 213, fm="pa")
f3o <- target.rot(f3)
f3o
```

The unweighted least squares solution may be shown graphically using the `fa.plot` function which is called by the generic plot function (Figure 4).  Factors were transformed obliquely using a `oblimin`.  These solutions may be shown as item by factor plots (Figure 4) or by a structure diagram (Figure 5).

```{r fig.asp=1}
plot(f3t)
```
Figure 4:  A graphic representation of the 3 oblique factors from the Thurstone data using
plot. Factors were transformed to an oblique solution using the oblimin function from the
GPArotation package.

#### Table 3:  
The 9 variable problem from Thurstone is a classic example of factoring where
there is a higher order factor, g, that accounts for the correlation between the factors.  The
factors were extracted using a weighted least squares algorithm.  All loadings are shown by
using the cut=0 option in the print.psych function.

```{r}
f3w <- fa(Thurstone, 3, n.obs = 213, fm="wls")
print(f3w, cut=0, digits=3)
```

```{r fig.asp=0.75}
fa.diagram(f3t)
```
Figure 5:  A graphic representation of the 3 oblique factors from the Thurstone data using
`fa.diagram`.  Factors were transformed to an oblique solution using `oblimin`.

A comparison of these three approaches suggests that the minres solution is more similar
to  a  maximum  likelihood  solution  and fits  slightly  better  than  the  `pa`  or  `wls`  solutions. Comparisons with SPSS suggest that the pa solution matches the SPSS OLS solution, but
that the minres solution is slightly better.  At least in one test data set, the weighted least
squares solutions,  although  tting equally well,  had slightly di erent structure loadings.
Note that the rotations used by SPSS will sometimes use the "Kaiser Normalization".  By
default, the rotations used in psych do not normalize, but this can be speci ed as an option
in fa.

