---
title: "R Notebook"
output: html_notebook
---

Source: http://stats.stackexchange.com/questions/121271/what-is-the-meaning-of-the-r-factanal-output


The chi-square statistic and p-value in factanal are testing the hypothesis that the model fits the data perfectly. When the p value is low, as it is here, we can reject this hypothesis - so in this case, the 2-factor model does not fit the data perfectly (this is opposite how it seems you were interpreting the output).

It's worth noting that 89.4% of the variance explained by two factors is very high, so I'm not sure why the 'only'.

The factors themselves are uncorrelated (orthogonal) but that doesn't mean individual measures cannot correlate with both factors. Think about the directions North and East on a compass - they're uncorrelated, but North-East would 'load' onto both of them positively.

Uniquenesses are the variance in each item that is not explained by the two factors.

This link might be useful to your interpretation.

```{r}
library(R.matlab)

# read the Matlab array
stockreturns.mat <- readMat("stockreturns.mat", fixNames = TRUE)
names(stockreturns.mat)

stocks <- stockreturns.mat$stocks
```



```{r}
factanal(stocks, 2)
```

